---
title: "ICONS Study: Bayesian approach to sample size and assurance"
output: rmarkdown::html_vignette
bibliography: references.bib
link-citations: yes
vignette: >
  %\VignetteIndexEntry{vignette-01}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette describes the use of functions within the `bayesiantrials` package to perform a Bayesian design and analysis of two-arm cluster randomised trials using assurance. This approach is described in @wilson2022. We will describe the use of the following functions:

- `SimulateGaussianCopula`
- `CalculateSampleSize`
- `CalculateAssurance`

We will apply the Bayesian approach to calculate an appropriate sample size for the ICONS cluster RCT @thomas2015. The trial considers the effectiveness of a systematic voiding programme for people admitted to NHS stroke units with urinary incontinence. The primary outcome is the incontinence symptom severity total score at 3 months post-randomisation. The voiding programme was compared to usual care. 

First we will load the package,
```{r setup}
library(bayesiantrials)
```

Set the number of samples and define the priors,

```{r eval=TRUE}
# Number of samples
N = 10000    

# Minimal clinically important difference delta
deltaMean = 3.5
deltaSd = 0.9 
deltaPrior = rnorm(N, deltaMean, deltaSd)

# Cluster size coefficient of variation nu
nuMean = 0.49
nuSd = 0.066
nuPrior = rgamma(N, (nuMean^2 / nuSd^2), (nuMean / nuSd^2)) 

# Overall mean effect lambda
lambdaMean = 1
lambdaSd = 0.001
lambdaPrior = rnorm(N, lambdaMean, lambdaSd)
```

We can load the ICC estimates directly from the package as `icons_icc_data`,

```{r eval=TRUE}
# Estimates of intra-cluster correlation rho
rho = icons_icc_data

# Standard deviation sigma
sigmaMean = 8.32
sigmaSd = 1
```

We define a joint prior distribution between $\rho$ and $\sigma$ via a Gaussian copula

```{r eval=TRUE}
# Correlation parameter for Gaussian Copula
gamma = 0.44

sim <- SimulateGaussianCopula(N, rho, sigmaMean, sigmaSd, gamma)

rhoPrior = sim[,1]
sigmaPrior = sim[,2]

sigmabPrior = sqrt(rhoPrior)*sigmaPrior
sigmawPrior = sigmaPrior*sqrt(1-rhoPrior)
```


Combine all priors into a matrix,

```{r eval=TRUE}
design = cbind(lambdaPrior, deltaPrior, nuPrior, sigmabPrior, sigmawPrior)
```


Calculate sample size,

```{r eval=TRUE}
# Standard value for posterior probability
prob = 0.05 

# Try with a cluster size of 45
clusterSize = 45

ss = CalculateSampleSize(1:50, design, L = 100, K = 1000, C = clusterSize, target = 0.8, alpha = prob)
ss*clusterSize

```



Check the assurance for this sample size,
```{r eval=TRUE}
CalculateAssurance(ss, design, L = 10, K = 1000, C = clusterSize)
CalculateAssurance(ss, design, L = 100, K = 1000, C = clusterSize)
CalculateAssurance(ss, design, L = 100, K = 1000, C = clusterSize)
```
  
  
The `CalculateSampleSize` and `CalculateAssurance` functions use a default model file for Bayesian inferences with `rjags`:

```{r comment = "", echo=FALSE}
cat(readLines("../inst/jags/mixed.R"), sep = "\n")
```

The user can also provide their own file for the JAGS model by specifying the relative path:

```{r eval = FALSE}
filename = "User/my_model_file.dat"

ss = CalculateSampleSize(1:50, design, L = 100, K = 1000, C = clusterSize, target = 0.8, modelFilePath = filename)

assure = CalculateAssurance(ss, design, L = 100, K = 1000, C = clusterSize, modelFilePath = filename)

```


### References


