---
title: "ICONS Study: Bayesian approach to sample size and assurance"
output: rmarkdown::html_vignette
bibliography: references.bib
link-citations: yes
vignette: >
  %\VignetteIndexEntry{vignette-01}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette describes the use of functions within the `bayesiantrials` package to perform a Bayesian design and analysis of two-arm cluster randomised trials using assurance. This approach is described in @wilson2022.

If not already installed, please follow the instructions on the package [GitHub page](https://github.com/NewcastleRSE/bayesiantrials) before loading the library.


```{r setup}
library(bayesiantrials)
```

## Assurance

## Sample Size

## ICONS Study

Set the number of samples,

```{r eval=TRUE}
# Number of samples
N = 10000    
```

Define priors.

```{r eval=TRUE}
# Minimal clinically important difference delta
m_delta = 3.5 # Mean of delta
s_delta = 0.9 # Stdev of delta

delta_pr = rnorm(N,m_delta,s_delta)

# Cluster size coefficient of variation
m_nu = 0.49    # Mean of nu
s_nu = 0.066   # Stdev of nu

nu_pr = rgamma(N,(m_nu^2/s_nu^2),(m_nu/s_nu^2)) 

# Overall mean effect lambda
m_lambda = 1     # Mean of lambda
s_lambda = 0.001 # Stdev of lambda

lambda_pr = rnorm(N,m_lambda,s_lambda)
```

We can load the ICC estimates directly from the package as `icons_icc_data`,

```{r eval=TRUE}
# Estimates of intra-cluster correlation rho
rho = icons_icc_data

# Standard deviation sigma
m_sigma = 8.32 # Mean of standard deviation sigma
s_sigma = 1    # Stdev of sigma
```

We define a joint prior distribution between $\rho$ and $\sigma$ via a Gaussian copula

```{r eval=TRUE}
# Correlation parameter for Gaussian Copula
gamma = 0.44

sim <- simulateGaussianCopula(N, rho, m_sigma, s_sigma, gamma)

rho_pr = sim[,1]
sigma_pr = sim[,2]

sigmab_pr = sqrt(rho_pr)*sigma_pr
sigmaw_pr = sigma_pr*sqrt(1-rho_pr)
```


Combine all priors into a matrix,

```{r eval=TRUE}
design = cbind(lambda_pr, delta_pr, nu_pr, sigmab_pr, sigmaw_pr)
```


Calculate sample size

```{r eval=TRUE}
# Standard value for posterior probability
prob = 0.05 

# Try with a cluster size of 45
clusterSize = 45

ss = getSampleSize(1:50, design, L = 100, K = 1000, C = clusterSize, target = 0.8, alpha = prob)
ss*clusterSize

```



Check the assurance for this sample size
```{r eval=TRUE}
getAssurance(ss, design, L = 10, K = 1000, C = clusterSize)
getAssurance(ss, design, L = 100, K = 1000, C = clusterSize)
getAssurance(ss, design, L = 100, K = 1000, C = clusterSize)
```
  
  

<!-- ### sample size calculation ### -->

<!-- ss.test = 10 -->
<!-- C.test = 40:50 -->

<!-- ### power ### -->

<!-- pow.ss = rep(0,length(C.test)) -->

<!-- for (i in 1:length(C.test)){ -->

<!-- pow.ss[i] = power.ss(100,delta,sigma,C=C.test[i],rho=median(rho),beta = beta,nu=nu, -->
<!--          tailed="one",cluster=TRUE,cluster.eq = FALSE) -->

<!-- } -->

<!-- pow.ss*C.test -->

<!-- ### hybrid MCID ### -->

<!-- mcid.ss = rep(0,length(C.test)) -->

<!-- for (i in 1:length(C.test)){ -->

<!-- mcid.ss[i] = assurance.ss(100,delta,sigma.pr,C=C.test[i],rho=rho.pr,nu = nu.pr,beta=beta, -->
<!--              cluster=TRUE,cluster.eq = FALSE,tailed="one") -->

<!-- } -->

<!-- mcid.ss*C.test -->

<!-- ### full hybrid ### -->

<!-- full.ss = rep(0,length(C.test)) -->

<!-- for (i in 1:length(C.test)){ -->

<!--   full.ss[i] = hybrid.ss(100,delta.pr,sigma.pr,C=C.test[i],rho=rho.pr,nu = nu.pr,beta=beta, -->
<!--                               cluster=TRUE,cluster.eq = FALSE,tailed="one") -->

<!-- } -->

<!-- full.ss*C.test -->

<!-- ### full Bayesian method ### -->

<!-- ss.bayes = rep(0,length(C.test)) -->

<!-- for (i in 1:length(C.test)){ -->

<!--   print(i) -->

<!--   ss.bayes[i] = sample.size(1:100,design,design.BL,N=5,C=C.test[i],target=0.8,100 -->
<!--                             ,method="INLA") -->

<!-- } -->

<!-- ss.bayes*C.test -->

<!-- ### changing the prior mean for delta ### -->

<!-- delta.mean = c(seq(5,2.5,-0.5),2.25) -->

<!-- sens.bayes = rep(0,length(delta.mean)) -->

<!-- for (j in 1:length(delta.mean)){ -->

<!--   print(j) -->

<!-- delta.sens = rnorm(N,delta.mean[j],0.9) -->

<!-- design.sens = cbind(alpha.pr,delta.sens,nu.pr,sigmab.pr,sigmaw.pr) -->

<!--   sens.bayes[j] = sample.size(1:100,design.sens,design.BL,N=5,C=C,target=(1-beta),100 -->
<!--                             ,method="INLA") -->

<!-- } -->

<!-- sens.bayes*40 -->

<!-- full.sens = rep(0,length(delta.mean)) -->

<!-- for (i in 1:length(delta.mean)){ -->

<!--   print(i) -->

<!--   delta.sens = rnorm(N,delta.mean[i],0.9) -->

<!--   full.sens[i] = hybrid.ss(200,delta.sens,sigma.pr,C=C,rho=rho.pr,nu = nu.pr,beta=beta, -->
<!--                          cluster=TRUE,cluster.eq = FALSE,tailed="one") -->

<!-- } -->

<!-- full.sens -->

<!-- plot(delta.mean,sens.bayes*C,type="l",col=2,xlab="Mean of delta",ylab="Sample size") -->
<!-- points(delta.mean,sens.bayes*C,col=2,pch=21,bg=2) -->
<!-- #lines(delta.mean,full.sens*C,col=4) -->
<!-- #points(delta.mean,full.sens*C,col=4,pch=21,bg=4) -->
<!-- abline(h=360,col=3,lty=2) -->
<!-- abline(h=480,col=5,lty=2) -->

<!-- #bayes.mean.data = data.frame(delta.mean,sens.bayes) -->
<!-- #colnames(bayes.mean.data)=c("mean","ss") -->

<!-- #write.table(bayes.mean.data,"SampleSize_Delta_Mean.txt") -->

<!-- ### changing the prior sd for delta ### -->

<!-- delta.sd = seq(0.8,2,0.2) -->

<!-- sd.bayes = rep(0,length(delta.sd)) -->

<!-- for (j in 2:2){ -->

<!--   print(j) -->

<!--   delta.sdev = rnorm(N,3.5,delta.sd[j]) -->

<!--   design.sens = cbind(alpha.pr,delta.sdev,nu.pr,sigmab.pr,sigmaw.pr) -->

<!--   sd.bayes[j] = sample.size(1:100,design.sens,design.BL,N=5,C=C,target=(1-beta),100 -->
<!--                               ,method="Bayes") -->

<!-- } -->

<!-- sd.bayes -->

<!-- full.sd = rep(0,length(delta.sd)) -->

<!-- for (i in 1:length(delta.sd)){ -->

<!--   print(i) -->

<!--   delta.sdev = rnorm(N,3.5,delta.sd[i]) -->

<!--   full.sd[i] = hybrid.ss(200,delta.sdev,sigma.pr,C=C,rho=rho.pr,nu = nu.pr,beta=beta, -->
<!--                            cluster=TRUE,cluster.eq = FALSE,tailed="one") -->

<!-- } -->

<!-- full.sd -->

<!-- plot(delta.sd[2:7],sd.bayes[2:7]*C,type="l",col=2,xlab="St Dev of delta",ylab="Sample size",ylim=c(200,600)) -->
<!-- points(delta.sd[2:7],sd.bayes[2:7]*C,col=2,pch=21,bg=2) -->
<!-- #lines(delta.sd,sd.bayes*C,col=4) -->
<!-- #points(delta.sd,sd.bayes*C,col=4,pch=21,bg=4) -->
<!-- abline(h=360,col=3,lty=2) -->
<!-- abline(h=480,col=5,lty=2) -->

<!-- #bayes.sd.data = data.frame(delta.sd,sd.bayes) -->
<!-- #colnames(bayes.sd.data)=c("sd","ss") -->

<!-- #write.table(bayes.sd.data,"SampleSize_Delta_SD.txt") -->

<!-- probs2 = 1-pnorm(2.52,3.5,delta.sd[2:7]) -->

<!-- plot(delta.sd[2:7],probs2,type="l",col=4,xlab="St Dev of delta",ylab="Probability") -->
<!-- points(delta.sd[2:7],probs2,col=4,pch=21,bg=4) -->



<!-- ``` -->

### References


